Development of a educational and sustainable robotic platform for children and adults

### 2014-03-06 doc/paper.tex
autocommited by makefile

\section{Introduction}
Getting children in touch with technology is important in todays technologically advanced society. In schools technology (specifically electrical engineering and programming) is often taught using robots. Robots are well suited for children since they allow interaction with the reality in contrast to arbitrary technology related exercises such as programming exercises which doesn't affect the real world.\newline 
Since schools are required to save money, robots used for teaching stay property of the schools. Therefore children usually cannot take the robot home to work with it. In addition modification of robot is usually not permitted which reduces potential learning experiences. Giving students the full control over all possibilities of the robot requires them to own the robot. This requires to robot to be cheap, extensible and suitable for educational purposes.\newline
The development of a robot platform for educational purpose requires requirement engineering, market research and evaluation of robots which are already used in education. Once the research is completed a new robot platform can be constructed taking the research results into account. The goal is to develop such an educational robot platform with regard to the target user group, their needs and educational concepts.
\section{Requirement Analysis / Project Goals}
User group of a robot can be categorized into different groups regardless of their age:
\begin{description}
\item[Beginner] \hfill \\ Beginners have little to no experience with robots. They need extensive assistance to perform tasks such as soldering and programming. For programming graphical programming languages (such as minibloq or ModKit) or languages providing a high level of abstraction (such as Wiring), 
\item[Advanced] \hfill \\ Advanced users have advanced experience with robots and robotics. Most programming and soldering tasks are done independently; working knowledge is utilized to solve new and unknown tasks.  
\item[Expert] \hfill \\ Expert users understand all aspects of robots and their internals. New challenges are faced with ease.
\end{description}

@@

The Asuro robot is developed by the German Aerospace Center (Deutsches Zentrum für Luft- und Raumfahrt e.V.) in cooperation with Arexx Engineering focused on educational use. Asuros development started in 2004 and is now discontinued. The robot is shipped as a do-it-yourself kit and requires about eight hours build time for a novice.

@@

\begin{table}[h!]
\centering
\begin{tabular}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{5}    & Having a price tag of 50 Euro renders the Asuro very affordable in comparison to the 100 Euro limit set. The cost effectiveness is outstanding and features sensors and actuators for basic tasks.     \\
Educational     & \stars{3}     & Asuro features detailed build-instructions in English and German. The guide also includes a tutorial on how to program Asuro in C. Even though there are a lot of opportunities to teach electrical engineering, the available documents do not explain the electrical usage and functional principle of each part. \\
Sustainable       & \stars{4}     & Asuro uses standard components only and features ROHS compliant parts which are free from heavy metals such as lead (PB). A broken Asuro can be easily repaired using the tools which are required to solder it. \\
Extendable & \stars{2}      & Having no free microcontroller pins and no broken out pins, Asuro cannot be considered extendable. There exist some hardware modification in the Asuro community which extend the functionality but require to unsolder/remove some parts.  \\
Trouble-free & \stars{3} & Building Asuro is well documented but some steps may lead to frustration. Mounting the axis using solder is very difficult and may burn the PCB due to the heat emitted in this process. Another frustrating task is to program the robot using the infrared serial connection, since this method is prone to interference. \\
Open-Source & \stars{4} & Asuro was licensed under the DLR-license which is transitioned into a Creative-Commons license model. Everything except the PCB layout files are available for download on the Asuro website.\\
\bottomrule
\end{tabular}
\caption{Asuro evaluation}
\label{tbl:asuro_eval}
\end{table}

@@

The Arduino Robot is the first robot created by the Arduino foundation. It comes preassembled and consists of two different boards: the motor board and the control board. Due to this design, the Arduino robot contains two Atmel ATmega 32U4 microcontroller featuring a clock speed of 16MHz, 32KB of flash memory (of which 28KB can be used for user programs), 2.5KB of RAM and 1KB of EEPROM each.

Each board can be individually programmed using the freely available Arduino IDE having the Robot connected to the computer using an USB cable. The IDE provides a lot of examples for the robot platform and introduces a library for abstract access to the robots functionaly. Power is applied to the system using four rechargeable NIMH-batteries which can be directly recharged on the board by connecting an appropriate power source to the board.

@@

\begin{table}[h!]
\centering
\begin{tabular}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{2}    & With a price tag of 189 Euro the Arduino Robot is far above the price limit of 100 Euro. The price for the basic robot is very high compared to its features.\\
Educational     & \stars{3}     & The Arduino robot comes pre-built and therefore lacks the build experience. Learning material for teachers is not provided but a lot of examples and a library to ease programming is provided. Instructions for other Arduino products can be easily adapted to the robot. \\
Sustainable       & \stars{4}     &  The robot integrates charging circuit for the NIMH batteries eliminates the need of a dedicated battery charger and therefore reduces the amount of electrical waste. A small restraint is the choice of components because they cannot be interchanged/repaired by beginners. This is due to the manufacturing process which requires SMD components to be cost effective. Through hole parts would be easier to interchange but require manual assembly which is very expensive.\\
Extendable & \stars{5}      &  Having prototyping areas directly on the control board, the robot can be considered extendable. Moreover connectors for additional sensors are populated all over the robot board making it possible and easy to add sensors without soldering.\\
Trouble-free & \stars{4} & Since the Arduino robot comes pre-assembled and tested in factory, the robot can be called trouble-free. Transportation defects like loose screws can be repaired using household items. The usage is very simple and straight-forward.\\
Open-Source & \stars{3} & Even though it is claimed that the Arduino robot is open-source, not all files required for building the robot are published. The website is missing the PCB design files and includes the schematics only. A software library is available under a open-source license. \\
\bottomrule
\end{tabular}
\caption{Arduino Robot evaluation}
\label{tbl:arduinorobot_eval}
\end{table}

To sum it up the Arduino robot is an extendable robot platform which requires no assembly at all. The integrated battery charging makes it a sustaina ble robot but the high price reduces the overall rating to an average of 3.5 stars.

@@

The LEGO Mindstorms NXT 2.0 platform was released by LEGO in 2009 and replaced the first LEGO Mindstorms NXT generation. It is compatible with Lego technic and therefore very flexible/extendable.

@@

\begin{table}[h!]
\centering
\begin{tabular}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{2}    & With a price tag of 279 Euro (retail consumer edition) and 368.99 Euro (education version) the robot is far beyond the price tag of 100 Euro. The amount of sensors and actuators is low compared to the price.\\
Educational & \stars{5}     & LEGO provides a lot of resources for teaching and personal education using the Mindstorms NXT 2.0. Even though the electrical components are not subject, the set can teach a lot of knowledge and even interdisciplinary content. The flexibility of the robot allows the transformation of the robert for a lot of education concern. \\
Sustainable  & \stars{3}     & The Mindstorms NXT 2.0 set cannot be repaired using household items. Nevertheless the components are considered sustainable to a certain extend, since they are designed to be robust and to last a long time.\\
Extendable & \stars{3} & The robot can be extended up to a certain extend. Even though the specification for each interface is public available custom sensors are rare and/or expensive. \\
Trouble-free & \stars{5} & Using the platform is trouble free and easy. The LEGO Mindstorms IDE is intuitional. Support for different programming languages such as Java exist. A huge community and a lot of resources are available.\\
Open-Source & \stars{4} & LEGO open sourced the LEGO Mindstorms NXT brick and moreover the specification for each connector and sensor. The PCB design files and the casing are not open-sourced to prevent unauthorized replicas.\\
\bottomrule
\end{tabular}
\caption{Lego Mindstorms evaluation}
\label{tbl:mindstorms_eval}
\end{table}

@@

\begin{table}[h!]
\centering
\begin{tabular}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{4}    & 99 CHF (without taxes directly from the Ecole Polytechnique Fédérale de Lausanne) is a reasonable price for a robot with this huge amount of features. On retail stores the price is higher due to profit margin.\\
Educational & \stars{3}     & Thymio II does not yet have a repository of educational material only, but it is planned to release such material. The compability with LEGO is a huge bonus, and creates a lot of pplications on which educational problems can be explained and demonstrated.\\
Sustainable  & \stars{3}     & The robot consists of small SMD parts and therefore parts cannot be replaced by the target user group. Materials used are suitable for children and designed to last. The integrated Li-Po battery can be recharged and therefore produces less environmental waste.\\
Extendable & \stars{3} & Except the LEGO compatible fixation points, Thymio II is not extendable without unsoldering components. On the software side, the choice of microcontroller supports a lot of additional program features.\\
Trouble-free & \stars{4} & Thymio II is trouble free since it comes preassembled and the usage is straight forward.\\
Open-Source & \stars{5} & Everything needed to build the robot is licensed under a permissive creative commons license. Program files and libraries are licensed under a LGPL license. Even the CAD files for the plastic mold injection parts are available without a charge.\\
\bottomrule
\end{tabular}
\caption{Thymio II evaluation}
\label{tbl:thymio_eval}
\end{table}

>>>>

\section{Introduction}
Getting children in touch with technology is important in todays technologically advanced society. In schools technology (specifically electrical engineering and programming) is often taught using robots. Robots are well suited for children since they allow interaction with the reality in contrast to arbitrary technology related exercises such as programming exercises which doesn't affect the real world.\newline 
Since schools are required to save money, robots used for teaching stay property of the schools. Therefore children usually cannot take the robot home to work with it. In addition modification of robot is usually not permitted which reduces potential learning experiences. Giving students the full control over all possibilities of the robot requires them to own the robot. This requires to robot to be cheap, extensible and suitable for educational purposes.\newline
The development of a robot platform for educational purpose requires requirement engineering, market research and evaluation of robots which are already used in education. Once the research is completed a new robot platform can be constructed taking the research results into account. The goal is to develop such an educational robot platform with regard to the target user group, their needs and educational concepts.
\section{Requirement Analysis / Project Goals}
User group of a robot can be categorized into different groups regardless of their age:
\begin{description}
\item[Beginner] \hfill \\ Beginners have little to no experience with robots. They need extensive assistance to perform tasks such as soldering and programming. For programming graphical programming languages (such as minibloq\footnote{\url{http://blog.minibloq.org/}} or ModKit\footnote{\url{http://www.modkit.com/}}) or languages providing a high level of abstraction (such as Wiring), 
\item[Advanced] \hfill \\ Advanced users have advanced experience with robots and robotics. Most programming and soldering tasks are done independently; working knowledge is utilized to solve new and unknown tasks.  
\item[Expert] \hfill \\ Expert users understand all aspects of robots and their internals. New challenges are faced with ease.
\end{description}

@@

The Asuro robot is developed by the German Aerospace Center (Deutsches Zentrum für Luft- und Raumfahrt e.V.) in cooperation with Arexx Engineering focused on educational use. Classroom courses, educational material and workshops were given using Asuro. Asuros development started in 2004 and is now discontinued. The robot is shipped as a do-it-yourself kit and requires about eight hours build time for a novice.

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{5}    & Having a price tag of 50 Euro renders the Asuro very affordable in comparison to the 100 Euro limit set. The cost effectiveness is outstanding and features sensors and actuators for basic tasks.     \\
Educational     & \stars{3}     & Asuro features detailed build-instructions in English and German. The guide also includes a tutorial on how to program Asuro in C. Even though there are a lot of opportunities to teach electrical engineering, the available documents do not explain the electrical usage and functional principle of each part. \\
Sustainable       & \stars{4}     & Asuro uses standard components only and features ROHS compliant parts which are free from heavy metals such as lead (PB). A broken Asuro can be easily repaired using the tools which are required to solder it. \\
Extendable & \stars{2}      & Having no free microcontroller pins and no broken out pins, Asuro cannot be considered extendable. There exist some hardware modification in the Asuro community which extend the functionality but require to unsolder/remove some parts.  \\
Trouble-free & \stars{3} & Building Asuro is well documented but some steps may lead to frustration. Mounting the axis using solder is very difficult and may burn the PCB due to the heat emitted in this process. Another frustrating task is to program the robot using the infrared serial connection, since this method is prone to interference. \\
Open-Source & \stars{4} & Asuro was licensed under the DLR-license which is transitioned into a Creative-Commons license model. Everything except the PCB layout files are available for download on the Asuro website.\\
\bottomrule
\caption{Asuro evaluation}
\label{tbl:asuro_eval}
\end{longtable}

@@

The Arduino Robot is the first robot created by the Arduino foundation targeted to educational purposes. It comes preassembled and consists of two different boards: the motor board and the control board. Due to this design, the Arduino robot contains two Atmel ATmega 32U4 microcontroller featuring a clock speed of 16MHz, 32KB of flash memory (of which 28KB can be used for user programs), 2.5KB of RAM and 1KB of EEPROM each.\footnote{{http://arduino.cc/en/Main/Robot}}

Each board can be individually programmed using the freely available Arduino IDE\footnote{\url{http://arduino.cc/en/Main/Software}} having the Robot connected to the computer using an USB cable. The IDE provides a lot of examples for the robot platform and introduces a library for abstract access to the robots functionaly. Power is applied to the system using four rechargeable NIMH-batteries which can be directly recharged on the board by connecting an appropriate power source to the board.

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable      & \stars{2}    & With a price tag of 189 Euro without VAT\footnote{As of 05.03.2014 from the official Arduino store: \url{http://store.arduino.cc/index.php?main_page=product_info&cPath=11_12&products_id=290}} the Arduino Robot is far above the price limit of 100 Euro. The price for the basic robot is very high compared to its features.\\
Educational     & \stars{3}     & The Arduino robot comes pre-built and therefore lacks the build experience. Learning material for teachers is not provided but a lot of examples and a library to ease programming is provided. Instructions for other Arduino products can be easily adapted to the robot. \\
Sustainable       & \stars{4}     &  The robot integrates charging circuit for the NIMH batteries eliminates the need of a dedicated battery charger and therefore reduces the amount of electrical waste. A small restraint is the choice of components because they cannot be interchanged/repaired by beginners. This is due to the manufacturing process which requires SMD components to be cost effective. Through hole parts would be easier to interchange but require manual assembly which is very expensive.\\
Extendable & \stars{5}      &  Having prototyping areas directly on the control board, the robot can be considered extendable. Moreover connectors for additional sensors are populated all over the robot board making it possible and easy to add sensors without soldering.\\
Trouble-free & \stars{4} & Since the Arduino robot comes pre-assembled and tested in factory, the robot can be called trouble-free. Transportation defects like loose screws can be repaired using household items. The usage is very simple and straight-forward.\\
Open-Source & \stars{3} & Even though it is claimed that the Arduino robot is open-source, not all files required for building the robot are published. The website is missing the PCB design files and includes the schematics only. A software library is available under a open-source license. \\
\bottomrule
\caption{Arduino Robot evaluation}
\label{tbl:arduinorobot_eval}
\end{longtable}

To sum it up the Arduino robot is an extendable robot platform which requires no assembly at all. The integrated battery charging makes it a sustainable robot but the high price reduces the overall rating to an average of 3.5 stars.

@@

The LEGO Mindstorms NXT 2.0 platform was released by LEGO in 2009 and replaced the first LEGO Mindstorms NXT generation. It is compatible with Lego technic and therefore very flexible/extendable and aims at educational (Education Kit) and personal use (Inventor Kit). It is one of the most successful robotics system in education and is in wide use.

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{2}    & With a price tag of 279 Euro (retail consumer edition) and 368.99 Euro (education version) the robot is far beyond the price tag of 100 Euro. The amount of sensors and actuators is low compared to the price.\\
Educational & \stars{5}     & LEGO provides a lot of resources for teaching and personal education using the Mindstorms NXT 2.0. Even though the electrical components are not subject, the set can teach a lot of knowledge and even interdisciplinary content. The flexibility of the robot allows the transformation of the robert for a lot of education concern. \\
Sustainable  & \stars{3}     & The Mindstorms NXT 2.0 set cannot be repaired using household items. Nevertheless the components are considered sustainable to a certain extend, since they are designed to be robust and to last a long time.\\
Extendable & \stars{3} & The robot can be extended up to a certain extend. Even though the specification for each interface is public available custom sensors are rare and/or expensive. \\
Trouble-free & \stars{5} & Using the platform is trouble free and easy. The LEGO Mindstorms IDE is intuitional. Support for different programming languages such as Java exist. A huge community and a lot of resources are available.\\
Open-Source & \stars{4} & LEGO open sourced the LEGO Mindstorms NXT brick and moreover the specification for each connector and sensor. The PCB design files and the casing are not open-sourced to prevent unauthorized replicas.\\
\bottomrule
\caption{Lego Mindstorms evaluation}
\label{tbl:mindstorms_eval}
\end{longtable}

@@

\begin{longtable}{p{0.2\textwidth}p{0.2\textwidth}p{0.6\textwidth}}
\toprule
Assessment Criteria    & Rating & Description \\
\midrule
Affordable  & \stars{4}    & 99 CHF (without taxes directly from the Ecole Polytechnique Fédérale de Lausanne) is a reasonable price for a robot with this huge amount of features. On retail stores the price is higher due to profit margin.\\
Educational & \stars{3}     & Thymio II does not yet have a repository of educational material only, but it is planned to release such material. The compability with LEGO is a huge bonus, and creates a lot of pplications on which educational problems can be explained and demonstrated.\\
Sustainable  & \stars{3}     & The robot consists of small SMD parts and therefore parts cannot be replaced by the target user group. Materials used are suitable for children and designed to last. The integrated Li-Po battery can be recharged and therefore produces less environmental waste.\\
Extendable & \stars{3} & Except the LEGO compatible fixation points, Thymio II is not extendable without unsoldering components. On the software side, the choice of microcontroller supports a lot of additional program features.\\
Trouble-free & \stars{4} & Thymio II is trouble free since it comes preassembled and the usage is straight forward.\\
Open-Source & \stars{5} & Everything needed to build the robot is licensed under a permissive creative commons license. Program files and libraries are licensed under a LGPL license. Even the CAD files for the plastic mold injection parts are available without a charge.\\
\bottomrule
\caption{Thymio II evaluation}
\label{tbl:thymio_eval}
\end{longtable}
